---
title: "Getting Started with Targets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Targets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The `wanikani` package provides a complete R interface to the WaniKani API v2. WaniKani is a Japanese learning platform that uses spaced repetition to teach kanji and vocabulary.

This vignette demonstrates basic usage. For pre-computed examples, we use the `targets` package to load cached results.

## Installation

```{r installation}
# Install from GitHub
devtools::install_github("johngavin/wanikani.r")
```

## Authentication

First, get your API token from WaniKani:
https://www.wanikani.com/settings/personal_access_tokens

```{r auth}
Sys.setenv(WANIKANI_API_TOKEN = "your_token_here")
```

## Load Package

```{r setup}
library(wanikani)
# Load pre-computed targets data if available
if (requireNamespace("targets", quietly = TRUE)) {
  library(targets)
}
```

## Basic Usage

### Get User Information

```{r user, eval=FALSE}
# Live API call (requires valid token)
user <- wk_user()
username <- user$data$username
current_level <- user$data$level
```

```{r user-show, eval=TRUE, echo=FALSE}
# Display pre-computed example
cat("User: example_user\n")
cat("Level: 5\n")
```

### Get Study Summary

```{r summary, eval=FALSE}
# Get upcoming lessons and reviews
summary <- wk_summary()
next_review <- summary$data$next_reviews_at
available_reviews <- length(summary$data$reviews)
```

## Working with Subjects

Subjects are the radicals, kanji, and vocabulary you learn.

### Get Subjects by Level

```{r subjects-level, eval=FALSE}
# Get all subjects from level 1
level1 <- wk_subjects(levels = 1)
num_items <- length(level1$data)
```

### Filter by Type

```{r subjects-type, eval=FALSE}
# Get only kanji from levels 1-3
kanji <- wk_subjects(
  types = "kanji",
  levels = c(1, 2, 3)
)

# Get vocabulary
vocabulary <- wk_subjects(types = "vocabulary", levels = 1)
```

### Get Specific Subject

```{r subjects-id, eval=FALSE}
# Get a specific subject by ID
subject <- wk_subjects_by_id(440)
print(subject$data$characters)  # "一"
print(subject$data$meanings)    # "one"
```

## Working with Assignments

Assignments represent your progress on specific subjects.

```{r assignments, eval=FALSE}
# Get all unlocked assignments
assignments <- wk_assignments(unlocked = TRUE)

# Get assignments at Apprentice stage (1-4)
apprentice <- wk_assignments(srs_stages = c(1, 2, 3, 4))

# Get assignments at Guru stage (5-6)
guru <- wk_assignments(srs_stages = c(5, 6))
```

## Reviews and Statistics

```{r reviews, eval=FALSE}
# Get recent reviews (last 7 days)
seven_days_ago <- format(Sys.time() - 7*24*60*60, "%Y-%m-%dT%H:%M:%SZ")
recent_reviews <- wk_reviews(updated_after = seven_days_ago)

# Get statistics for items you struggle with
struggling <- wk_review_statistics(percentages_less_than = 75)
```

## Study Materials

Add custom notes and synonyms to subjects.

```{r study-materials, eval=FALSE}
# Add a custom note and synonyms
material <- wk_study_materials_create(
  subject_id = 440,
  meaning_note = "Think of a tall tree reaching to heaven",
  meaning_synonyms = c("tree", "wood", "timber")
)

# Update existing study material
updated <- wk_study_materials_update(
  id = material$data$id,
  meaning_note = "Updated: Looks like a tree with branches"
)
```

## Pagination

Many endpoints return paginated results.

```{r pagination, eval=FALSE}
# Fetch all pages automatically
all_subjects <- wk_fetch_all_pages(
  wk_subjects,
  types = "kanji",
  verbose = TRUE
)

cat(sprintf("Fetched %d kanji across %d pages\n",
            all_subjects$total_count,
            all_subjects$page_count))
```

## Advanced Usage: R6 Client

For more control, use the `WaniKaniClient` R6 class directly.

```{r r6-client, eval=FALSE}
# Create a client
client <- WaniKaniClient$new(api_token = "your_token")

# Make custom requests
response <- client$get("/subjects", query = list(
  levels = "1,2,3",
  types = "kanji"
))

# Check rate limit status
remaining <- client$rate_limit_remaining
reset_time <- as.POSIXct(client$rate_limit_reset, origin = "1970-01-01")

cat(sprintf("Remaining requests: %d\nResets at: %s\n",
            remaining, reset_time))
```

## Using Pre-computed Data with Targets

If you're building vignettes and have pre-computed data:

```{r targets-load, eval=FALSE}
# Load pre-computed targets
if (file.exists("_targets/meta/meta")) {
  targets::tar_load(example_user_data)
  targets::tar_load(example_subjects)

  # Use the pre-computed data
  print(example_user_data$data$username)
  print(example_subjects$data[[1]]$data$characters)
}
```

## Rate Limiting

The WaniKani API has a rate limit of 60 requests per minute.

```{r rate-limit, eval=FALSE}
client <- WaniKaniClient$new()

for (i in 1:5) {
  response <- client$get("/user")
  cat(sprintf("Request %d - Remaining: %d\n", i, client$rate_limit_remaining))

  # Wait if approaching limit
  if (client$rate_limit_remaining < 5) {
    wait_until <- as.POSIXct(client$rate_limit_reset, origin = "1970-01-01")
    wait_seconds <- as.numeric(difftime(wait_until, Sys.time(), units = "secs"))
    message(sprintf("Approaching rate limit. Waiting %d seconds...", wait_seconds))
    Sys.sleep(wait_seconds)
  }
}
```

## Conclusion

For more details on specific functions, see their help pages:

```{r help, eval=FALSE}
?wk_subjects
?wk_assignments
?WaniKaniClient
```

Happy studying!
