---
title: "Targets Pipeline Visualization and Diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Targets Pipeline Visualization and Diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Check if targets is available - if not, skip all chunk evaluation
has_targets <- requireNamespace("targets", quietly = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = FALSE,
  eval = has_targets  # Only evaluate chunks if targets is available
)
```

## Introduction

This vignette provides comprehensive visualization and diagnostics of the `targets` pipeline used to pre-compute data for the wanikani package vignettes. It demonstrates:

- Pipeline structure and dependencies
- Build times and performance metrics
- Memory usage statistics
- Target status and metadata
- Session information

```{r check-targets, echo=FALSE, results='asis', eval=TRUE}
if (!requireNamespace("targets", quietly = TRUE)) {
  cat("**Note:** This vignette requires the `targets` package to be installed. ")
  cat("Install it with `install.packages('targets')` to view the full content.\n\n")
}
```

## Setup

```{r setup}
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)

# Check if pipeline has been run
pipeline_exists <- file.exists("_targets/meta/meta")

if (!pipeline_exists) {
  message("Pipeline has not been run yet. Run targets::tar_make() first.")
  message("For this vignette, we'll show the expected structure.")
}
```

## Pipeline Structure

### Dependency Graph

The targets pipeline shows the relationships between different data objects:

```{r network-plot, eval=pipeline_exists, fig.height=7}
# Interactive network visualization
if (requireNamespace("visNetwork", quietly = TRUE)) {
  targets::tar_visnetwork(
    targets_only = TRUE,
    label = c("time", "size", "branches")
  )
} else {
  message("Install visNetwork for interactive pipeline visualization:")
  message("install.packages('visNetwork')")
}
```

### Static Dependency Graph

```{r glimpse-plot, eval=pipeline_exists}
# Static visualization
targets::tar_glimpse()
```

### Target Manifest

Complete list of all targets in the pipeline:

```{r manifest}
if (pipeline_exists) {
  manifest <- targets::tar_manifest()
  knitr::kable(
    manifest,
    caption = "Pipeline Manifest: All Targets",
    format = "markdown"
  )
}
```

## Performance Metrics

### Build Time Analysis

```{r build-times, eval=pipeline_exists, fig.height=5}
# Get build time metadata
meta <- targets::tar_meta(fields = c("name", "time", "bytes", "seconds"))

if (nrow(meta) > 0) {
  # Convert to data frame using tidyverse
  meta_df <- meta %>%
    as.data.frame() %>%
    mutate(
      seconds = as.numeric(seconds),
      time_formatted = format(time, "%Y-%m-%d %H:%M:%S")
    )

  # Plot build times
  meta_df %>%
    ggplot(aes(x = reorder(name, seconds), y = seconds)) +
    geom_col(fill = "steelblue") +
    coord_flip() +
    labs(
      title = "Target Build Times",
      subtitle = "Time taken to build each target",
      x = "Target Name",
      y = "Time (seconds)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      axis.text.y = element_text(size = 10)
    )
}
```

### Build Time Statistics

```{r time-stats, eval=pipeline_exists}
if (pipeline_exists && nrow(meta) > 0) {
  # Use tidyverse to calculate statistics
  time_summary <- meta_df %>%
    summarise(
      total = sum(seconds, na.rm = TRUE),
      mean = mean(seconds, na.rm = TRUE),
      median = median(seconds, na.rm = TRUE),
      max_val = max(seconds, na.rm = TRUE),
      min_val = min(seconds, na.rm = TRUE)
    )

  max_target <- meta_df %>% slice_max(seconds, n = 1) %>% pull(name)
  min_target <- meta_df %>% slice_min(seconds, n = 1) %>% pull(name)

  time_stats <- tibble(
    Metric = c("Total Time", "Mean Time", "Median Time", "Max Time", "Min Time"),
    Value = c(
      sprintf("%.2f seconds", time_summary$total),
      sprintf("%.2f seconds", time_summary$mean),
      sprintf("%.2f seconds", time_summary$median),
      sprintf("%.2f seconds (%s)", time_summary$max_val, max_target),
      sprintf("%.2f seconds (%s)", time_summary$min_val, min_target)
    )
  )

  knitr::kable(
    time_stats,
    caption = "Build Time Statistics",
    format = "markdown",
    align = c("l", "r")
  )
}
```

## Memory Usage

### Memory by Target

```{r memory-usage, eval=pipeline_exists, fig.height=5}
if (pipeline_exists && nrow(meta) > 0) {
  # Convert bytes to MB using tidyverse
  meta_df <- meta_df %>%
    mutate(mb = bytes / (1024^2))

  # Plot memory usage
  meta_df %>%
    ggplot(aes(x = reorder(name, mb), y = mb)) +
    geom_col(fill = "coral") +
    coord_flip() +
    labs(
      title = "Target Memory Usage",
      subtitle = "Size of each target in memory",
      x = "Target Name",
      y = "Size (MB)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      axis.text.y = element_text(size = 10)
    )
}
```

### Memory Statistics

```{r memory-stats, eval=pipeline_exists}
if (pipeline_exists && nrow(meta) > 0) {
  memory_stats <- data.frame(
    Metric = c("Total Size", "Mean Size", "Median Size", "Largest", "Smallest"),
    Value = c(
      sprintf("%.2f MB", sum(meta_df$mb, na.rm = TRUE)),
      sprintf("%.2f MB", mean(meta_df$mb, na.rm = TRUE)),
      sprintf("%.2f MB", median(meta_df$mb, na.rm = TRUE)),
      sprintf("%.2f MB (%s)", max(meta_df$mb, na.rm = TRUE),
              meta_df$name[which.max(meta_df$mb)]),
      sprintf("%.2f MB (%s)", min(meta_df$mb, na.rm = TRUE),
              meta_df$name[which.min(meta_df$mb)])
    )
  )

  knitr::kable(
    memory_stats,
    caption = "Memory Usage Statistics",
    format = "markdown",
    align = c("l", "r")
  )
}
```

### Memory vs. Build Time

```{r memory-time-scatter, eval=pipeline_exists, fig.height=5}
if (pipeline_exists && nrow(meta) > 0) {
  ggplot(meta_df, aes(x = seconds, y = mb)) +
    geom_point(size = 3, color = "darkgreen", alpha = 0.7) +
    geom_text(aes(label = name), vjust = -0.5, size = 3) +
    labs(
      title = "Memory Usage vs Build Time",
      subtitle = "Relationship between target size and build duration",
      x = "Build Time (seconds)",
      y = "Memory (MB)"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold", size = 14))
}
```

## Pipeline Status

### Target Progress

```{r progress, eval=pipeline_exists}
if (pipeline_exists) {
  progress <- targets::tar_progress()

  if (nrow(progress) > 0) {
    # Summary of target status
    status_summary <- table(progress$progress)

    status_df <- data.frame(
      Status = names(status_summary),
      Count = as.integer(status_summary)
    )

    knitr::kable(
      status_df,
      caption = "Target Status Summary",
      format = "markdown",
      col.names = c("Status", "Number of Targets")
    )
  }
}
```

### Outdated Targets

```{r outdated, eval=pipeline_exists}
if (pipeline_exists) {
  outdated <- targets::tar_outdated()

  if (length(outdated) > 0) {
    cat("The following targets are outdated and need rebuilding:\n")
    cat(paste("-", outdated, collapse = "\n"))
  } else {
    cat("✓ All targets are up to date!\n")
  }
}
```

## Target Metadata

### Complete Metadata Table

```{r metadata-table, eval=pipeline_exists}
if (pipeline_exists && nrow(meta) > 0) {
  # Create readable metadata table
  meta_table <- data.frame(
    Target = meta_df$name,
    `Build Time` = sprintf("%.3f sec", meta_df$seconds),
    `Size` = sprintf("%.2f MB", meta_df$mb),
    `Last Built` = meta_df$time_formatted,
    stringsAsFactors = FALSE
  )

  knitr::kable(
    meta_table,
    caption = "Complete Target Metadata",
    format = "markdown",
    align = c("l", "r", "r", "l")
  )
}
```

### Warnings and Errors

```{r warnings-errors, eval=pipeline_exists}
if (pipeline_exists) {
  meta_full <- targets::tar_meta(fields = c("name", "warnings", "error"))

  # Check for warnings
  warnings_present <- sapply(meta_full$warnings, function(x) !is.null(x) && length(x) > 0)

  if (any(warnings_present)) {
    cat("⚠ Warnings detected in the following targets:\n")
    for (i in which(warnings_present)) {
      cat(sprintf("\n%s:\n", meta_full$name[i]))
      cat(paste("  ", meta_full$warnings[[i]], collapse = "\n"))
    }
  } else {
    cat("✓ No warnings in any targets\n")
  }

  # Check for errors
  errors_present <- sapply(meta_full$error, function(x) !is.null(x) && nchar(x) > 0)

  if (any(errors_present)) {
    cat("\n\n❌ Errors detected in the following targets:\n")
    for (i in which(errors_present)) {
      cat(sprintf("\n%s:\n", meta_full$name[i]))
      cat(paste("  ", meta_full$error[i]))
    }
  } else {
    cat("\n✓ No errors in any targets\n")
  }
}
```

## Dependency Analysis

### Target Dependencies

```{r dependencies, eval=pipeline_exists}
if (pipeline_exists) {
  # Get network data
  network <- targets::tar_network(targets_only = TRUE)

  if (!is.null(network$edges) && nrow(network$edges) > 0) {
    # Create dependency table
    deps_table <- data.frame(
      From = network$edges$from,
      To = network$edges$to,
      stringsAsFactors = FALSE
    )

    knitr::kable(
      deps_table,
      caption = "Target Dependencies",
      format = "markdown"
    )
  } else {
    cat("No dependencies between targets (all independent)\n")
  }
}
```

### Dependency Depth

```{r dep-depth, eval=pipeline_exists, fig.height=4}
if (pipeline_exists) {
  # Calculate dependency depth for each target
  network <- targets::tar_network(targets_only = TRUE)

  if (!is.null(network$vertices) && nrow(network$vertices) > 0) {
    depth_df <- data.frame(
      target = network$vertices$name,
      depth = network$vertices$level
    )

    ggplot(depth_df, aes(x = factor(depth), fill = factor(depth))) +
      geom_bar() +
      labs(
        title = "Pipeline Depth Distribution",
        subtitle = "Number of targets at each dependency level",
        x = "Dependency Level",
        y = "Number of Targets"
      ) +
      scale_fill_brewer(palette = "Set3") +
      theme_minimal() +
      theme(
        plot.title = element_text(face = "bold", size = 14),
        legend.position = "none"
      )
  }
}
```

## System Information

### R Session Info

```{r session-info}
# Get session information
sess_info <- sessionInfo()

# Create formatted output
cat("R Version:", sess_info$R.version$version.string, "\n")
cat("Platform:", sess_info$platform, "\n")
cat("Running under:", sess_info$running, "\n\n")

cat("Locale:\n")
cat(" ", sess_info$locale, "\n\n")

cat("Loaded Packages:\n")
loaded_pkgs <- sess_info$loadedOnly
if (length(loaded_pkgs) > 0) {
  for (pkg_name in names(loaded_pkgs)) {
    pkg <- loaded_pkgs[[pkg_name]]
    cat(sprintf("  %s %s\n", pkg_name, pkg$Version))
  }
}
```

### Package Versions Table

```{r package-versions}
# Key packages for this pipeline
key_packages <- c("targets", "tarchetypes", "wanikani", "ggplot2", "knitr", "rmarkdown")

versions <- data.frame(
  Package = key_packages,
  Version = sapply(key_packages, function(pkg) {
    if (requireNamespace(pkg, quietly = TRUE)) {
      as.character(packageVersion(pkg))
    } else {
      "Not installed"
    }
  }),
  stringsAsFactors = FALSE
)

knitr::kable(
  versions,
  caption = "Key Package Versions",
  format = "markdown",
  row.names = FALSE
)
```

### System Resources

```{r system-resources}
# Get system information
sys_info <- Sys.info()

system_df <- data.frame(
  Property = c("System", "Release", "Version", "Machine", "User"),
  Value = c(
    sys_info["sysname"],
    sys_info["release"],
    sys_info["version"],
    sys_info["machine"],
    sys_info["user"]
  ),
  stringsAsFactors = FALSE
)

knitr::kable(
  system_df,
  caption = "System Information",
  format = "markdown",
  row.names = FALSE
)
```

### Pipeline Environment

```{r pipeline-env, eval=pipeline_exists}
if (pipeline_exists) {
  # Try to load pipeline metadata target
  tryCatch({
    targets::tar_load(pipeline_metadata)

    env_df <- data.frame(
      Property = c(
        "Pipeline Name",
        "Created At",
        "R Version",
        "Platform",
        "Targets Version"
      ),
      Value = c(
        pipeline_metadata$pipeline_name,
        as.character(pipeline_metadata$created_at),
        pipeline_metadata$r_version,
        pipeline_metadata$platform,
        as.character(pipeline_metadata$targets_version)
      ),
      stringsAsFactors = FALSE
    )

    knitr::kable(
      env_df,
      caption = "Pipeline Environment Details",
      format = "markdown",
      row.names = FALSE
    )
  }, error = function(e) {
    cat("Pipeline metadata not available\n")
  })
}
```

## Performance Summary

### Overall Pipeline Statistics

```{r overall-stats, eval=pipeline_exists}
if (pipeline_exists && nrow(meta) > 0) {
  summary_stats <- data.frame(
    Metric = c(
      "Total Targets",
      "Total Build Time",
      "Total Memory Used",
      "Average Build Time per Target",
      "Average Memory per Target",
      "Fastest Target",
      "Slowest Target",
      "Smallest Target",
      "Largest Target"
    ),
    Value = c(
      nrow(meta_df),
      sprintf("%.2f seconds", sum(meta_df$seconds, na.rm = TRUE)),
      sprintf("%.2f MB", sum(meta_df$mb, na.rm = TRUE)),
      sprintf("%.2f seconds", mean(meta_df$seconds, na.rm = TRUE)),
      sprintf("%.2f MB", mean(meta_df$mb, na.rm = TRUE)),
      sprintf("%s (%.2f sec)", meta_df$name[which.min(meta_df$seconds)],
              min(meta_df$seconds, na.rm = TRUE)),
      sprintf("%s (%.2f sec)", meta_df$name[which.max(meta_df$seconds)],
              max(meta_df$seconds, na.rm = TRUE)),
      sprintf("%s (%.2f MB)", meta_df$name[which.min(meta_df$mb)],
              min(meta_df$mb, na.rm = TRUE)),
      sprintf("%s (%.2f MB)", meta_df$name[which.max(meta_df$mb)],
              max(meta_df$mb, na.rm = TRUE))
    ),
    stringsAsFactors = FALSE
  )

  knitr::kable(
    summary_stats,
    caption = "Overall Pipeline Performance Summary",
    format = "markdown",
    align = c("l", "r")
  )
}
```

## Recommendations

Based on the pipeline analysis:

```{r recommendations, eval=pipeline_exists, results='asis'}
if (pipeline_exists && nrow(meta) > 0) {
  cat("\n### Performance Recommendations\n\n")

  # Check for slow targets
  slow_threshold <- quantile(meta_df$seconds, 0.75)
  slow_targets <- meta_df$name[meta_df$seconds > slow_threshold]

  if (length(slow_targets) > 0) {
    cat("- **Slow targets detected**: Consider optimizing:\n")
    for (target in slow_targets) {
      cat(sprintf("  - `%s` (%.2f seconds)\n",
                  target,
                  meta_df$seconds[meta_df$name == target]))
    }
    cat("\n")
  }

  # Check for large targets
  large_threshold <- quantile(meta_df$mb, 0.75)
  large_targets <- meta_df$name[meta_df$mb > large_threshold]

  if (length(large_targets) > 0) {
    cat("- **Large targets detected**: Consider data reduction for:\n")
    for (target in large_targets) {
      cat(sprintf("  - `%s` (%.2f MB)\n",
                  target,
                  meta_df$mb[meta_df$name == target]))
    }
    cat("\n")
  }

  cat("- **Pipeline health**: ✓ All targets built successfully\n")
  cat("- **Best practices**: Use `targets::tar_make()` to rebuild only outdated targets\n")
  cat("- **Caching**: Pipeline uses intelligent caching - only changed targets are rebuilt\n")
}
```

## Conclusion

This vignette provides comprehensive diagnostics for the targets pipeline. Use these visualizations and statistics to:

- Monitor pipeline performance
- Identify bottlenecks
- Optimize slow or large targets
- Track changes over time
- Debug pipeline issues

For more information on targets, see the [targets documentation](https://docs.ropensci.org/targets/).
